\bf The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth

### 2021-09-16 paper/paper-jsdse/paper-jsdse.tex

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

@@

\if0\blind
{
  \title{\bf The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth}
  \author{Author 1\thanks{
    The authors gratefully acknowledge \textit{please remember to list all relevant funding sources in the unblinded version}}\hspace{.2cm}\\
    Department of YYY, University of XXX\\
    and \\
    Author 2 \\
    Department of ZZZ, University of WWW}
  \maketitle
} \fi

@@

NSLY79 is a prominent open data source that has been an important support for educational purposes and multidisciplinary research on longitudinal data. Subsets of this data can be found in numerous textbooks and research articles. However, the steps and decisions taken to get from the original data to these subsets is never clearly articulated. This article describes our journey when trying to re-create a textbook example data set from the original database. Ultimately, the goal is to provide the resources so that refreshing data subsets can be done on a regular basis, especially when updates are made to the database. Thus, this paper demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research. A brief exploratory analysis of the resulting data is also conducted. Three new data sets, and the code to produce them are provided in an accompanying open source R package, called \texttt{yowie}. As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.

\textbf{Keywords}: Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression

\hypertarget{intro}{%
\section{Introduction}\label{intro}}

``Open data'' is data that is freely accessible, modifiable, and shareable by anyone for any purpose (Open Knowledge Foundation 2021). This type of data can be useful as example data in statistical textbooks and for research purposes. However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape. Huebner Marianne, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative. Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (M. Huebner et al. 2020).

Data cleaning is part of what is called ``initial data analysis (IDA)'' (Chatfield 1985). The other IDA steps are to explore the data properties and check the assumption of modeling to document and report the process for the later formal analysis. Hence, exploratory data analysis, which is defined by Tukey (1977) as a detective work to get the clue from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA. Dasu and Johnson (2003) say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80\% of the data mining task.

Despite its importance, this IDA stage is often undervalued and neglected (Chatfield 1985). There are few research papers that document the data cleaning (Wickham 2014). Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (M. Huebner et al. 2020).

This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NSLY79. This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century (Pergamit et al. 2001). In addition, The National Longitudinal Survey is considered a survey with high retention rates and carefully designed, making it suitable for a life course research (Pergamit et al. 2001) and (Cooksey 2017). According to Cooksey (2017), thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data. Moreover, the NLSY79 is considered the most widely used and most important cohort in the NLSY79 data sets (Pergamit et al. 2001).

Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling. Our aim is to refresh this textbook data to be more up to date. Here, we investigate the process of getting from the raw NLSY79 to this textbook data set. However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.

This paper is structured to have 5 sections. Section \ref{database} describes the original data source. Section \ref{cleaning} presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and initial data analysis to find and treat the anomalies in the data. We also gave examples of exploratory data analysis using the clean data in Section \ref{eda}. Finally, Section \ref{summary} summarises the paper.

\hypertarget{database}{%
\section{The NLSY79}\label{database}}

\hypertarget{database-1}{%
\subsection{Database}\label{database-1}}

The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth and born between 1957-1964 (The U.S. Bureau of Labor Statistics, n.d.). The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979. It was comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military. In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects are the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics, respectively. Hence, 9,964 respondents remain in the eligible samples. The surveys were conducted annually from 1979 to 1994 and biennially thereafter. Data are now available from Round 1 (1979 survey year) to Round 28 (2018 survey year).

Although the main focus area of the NLSY is labor and employment, the NLSY also covers several other topics, including education; training and achievement; household, geography, and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.

There are two ways to conduct the interview of the NLSY79, which are face-to-face interviews or by telephone. In recent survey years, more than 90 percent of respondents were interviewed by telephone (Cooksey 2017).

\hypertarget{target-data}{%
\subsection{Target Data}\label{target-data}}

As mentioned in Section \ref{intro}, this paper aims to refresh the NLSY79 data as used in Singer and Willet's textbook (Singer and Willett (2003)). This data set contains the information of yearly mean hourly wages from the NLSY79 cohort, with education and race as covariates. This data set represents the measurement of male high-school dropouts, aged from 14 to 17 years old when first time measured. Accordingly, to get this data, we have two additional target data, NSLY79 cohort's demographic profiles and wages data of people who completed education up to high school.

Since there is no information on high school dropouts in Singer and Willet's book, we define high school dropouts as respondents who only completed either \(9^{th}\), \(10^{th}\), \(11^{th}\) grade, or who completed \(12^{th}\) when their age is at least 19 years old (means that these IDs being dropped out in a certain year, but they returned back to high school to complete it).

\hypertarget{cleaning}{%
\section{The NLSY79 Data Cleaning}\label{cleaning}}

\hypertarget{getting-and-tidying-the-data}{%
\subsection{Getting and Tidying the Data}\label{getting-and-tidying-the-data}}

The NLSY79 data are stored in a \href{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}{database} and could be downloaded by variables. Several variables are available for download and could be browsed by index. For the wages data set, we only extracted these variables:

\textbf{Education, Training \& Achievement Scores}

\begin{itemize}
\tightlist
\item
  Education -\textgreater{} Summary measures -\textgreater{} All schools -\textgreater{} By year -\textgreater{} Highest grade completed

  \begin{itemize}
  \tightlist
  \item
    Downloaded all of variables in Highest grade completed.
  \end{itemize}
\end{itemize}

\textbf{Employment}

\begin{itemize}
\tightlist
\item
  Summary measures -\textgreater{} By job -\textgreater{} Hours worked and Hourly wages

  \begin{itemize}
  \tightlist
  \item
    Downloaded all of the variables in Hours worked.
  \item
    Downloaded all of the variables in Hourly wages
    Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
  \end{itemize}
\end{itemize}

\textbf{Household, Geography \& Contextual Variables}

\begin{itemize}
\tightlist
\item
  Context -\textgreater{} Summary measures -\textgreater{} Basic demographics

  \begin{itemize}
  \tightlist
  \item
    Downloaded year and month of birth, race, and sex variables.
    There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
  \end{itemize}
\end{itemize}

The downloaded data set came in a csv (NLSY79.csv) and dat (NLSY79.dat) format. We only used the .dat format. It also came along with these files:

\begin{itemize}
\tightlist
\item
  NLSY79.NLSY79: Tagset of variables that can be uploaded to the website to recreate the data set.
\item
  NLSY79.R: R script provided automatically by the database for reading the data into R and converting the variables' names and label into something more sensible. We utilized this code to do an initial data tidying. It produced two data set, \texttt{categories\_qnames} (the observations are stored in categorical/interval values) and \texttt{new\_data\_qnames} (the observations are stored in integer form).
\end{itemize}

According to Wickham (2014), tidy data sets comply with three rules. Firstly, each variable forms a column. Secondly, each observation forms a row. Lastly, each type of observational unit forms a table. Unfortunately, the \texttt{new\_data\_qnames} did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable. Hence, the data contains a huge number of columns (686 columns). The example of the untidy data set is displayed in Table \ref{tab:untidy-data}. The table intended to display the hourly rate of each respondent by job (HRP1 to HRP5) and by year (1979 and 1980). The table implies that the column headers are values of the year and job, not variable names. Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables we want to put in the final data set. We mainly used \texttt{tidyr} (Wickham 2020b), \texttt{dplyr} (Wickham et al. 2020), and \texttt{stringr} (Wickham 2019) to do this job.

\begin{table}

\caption{\label{tab:untidy-data}The Untidy Form of the NLSY79 Raw Data}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r}
\hline
CASEID\_1979 & HRP1\_1979 & HRP2\_1979 & HRP3\_1979 & HRP4\_1979 & HRP5\_1979 & HRP1\_1980\\
\hline
1 & 328 & NA & NA & NA & NA & NA\\
\hline
2 & 385 & NA & NA & NA & NA & 457\\
\hline
3 & 365 & NA & 275 & NA & NA & 397\\
\hline
4 & NA & NA & NA & NA & NA & NA\\
\hline
5 & 310 & 375 & NA & NA & NA & 333\\
\hline
6 & NA & NA & NA & 250 & NA & 275\\
\hline
\end{tabular}
\end{table}

\hypertarget{tidying-demographic-variables}{%
\subsubsection{Tidying demographic variables}\label{tidying-demographic-variables}}

\texttt{Age\ in\ 1979}, \texttt{gender}, \texttt{race}, \texttt{highest\ grade\ completed}, and the \texttt{year\ when\ the\ highest\ grade\ completed} are the variables that we want to use in the cleaned data set.

Age in 1979 are derived from year and month of birth variables in the raw data. The variables have two versions, which are the 1979 version and the 1981 version. We only used the 1979 data and did a consistency check of those years and flag the inconsistent observations.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#tidy the date of birth data}
\NormalTok{dob\_tidy }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
                \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"Q1{-}3\_A\textasciitilde{}"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{dob\_year =} \FunctionTok{case\_when}\NormalTok{(}
    \CommentTok{\# if the years recorded in both sets match, take 79 data}
    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
    \CommentTok{\# if the year in the 81 set is missing, take the 79 data}
    \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
    \CommentTok{\# if the sets don\textquotesingle{}t match for dob year, take the 79 data}
    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
    \AttributeTok{dob\_month =} \FunctionTok{case\_when}\NormalTok{(}
      \CommentTok{\# if months recorded in both sets match, take 79 data}
      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
      \CommentTok{\# if month in 81 set is missing, take the 79 data}
      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
      \CommentTok{\# if sets don\textquotesingle{}t match for dob month, take the 79 data}
      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
    \CommentTok{\# flag if there is a conflict between dob recorded in 79 and 81}
    \AttributeTok{dob\_conflict =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
\NormalTok{      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
\NormalTok{      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&}
\NormalTok{        (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{,}
      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{|} \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
\NormalTok{                dob\_month,}
\NormalTok{                dob\_year,}
\NormalTok{                dob\_conflict)}
\end{Highlighting}
\end{Shaded}

For \texttt{gender} and \texttt{race}, we only rename the column names. It is worth noting that using these two variable needs special attention. Gender, as reported in the data, only has two categories, which is recognised today as inadequate. Gender is not binary. Further, race is as reported in the database. When doing analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.

The next step is tidying the \texttt{highest\ grade\ completed} variable. This variable came with several version in each year. We choose the revised May data because it seemed to have less missing and presumably has been checked. However, there is no revised May data for 2012, 2014, 2016, and 2018. Thus, we just use the ordinary May data for these years. The code for tidying this variable is provided in the supplementary materials.

Further, \texttt{highest\ grade\ completed} is measured and could be updated in each period of the survey. Hence, we only used the highest grade completed ever and derived the year when it is completed. The code for tidying this variable is also available in the supplementary materials.

Finally, we get all of the demographic profiles of the NLSY79 cohort. We then save this data as \texttt{demog\_nlsy79}.

\hypertarget{tidying-employment-variables}{%
\subsubsection{Tidying employment variables}\label{tidying-employment-variables}}

The employment data comprises of three variables, i.e., \texttt{total\ hours\ of\ work\ per\ week}, \texttt{number\ of\ jobs\ that\ an\ individual\ has}, and \texttt{mean\ hourly\ wage}.

\texttt{hours\ worked\ per\ week} initially has only one version per job, no choice from 1979 to 1987 (QES-52A). From 1988 onward, when we had more options, we chose the variable for total hours, including time spent working from home (QES-52D). However, in 1993, this variable did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead. In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we use these.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# make a list for years where we used the "QES{-}52A"}
\NormalTok{year\_A }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
\CommentTok{\#function to get the hour of work}
\NormalTok{get\_hour }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year)\{}
  \ControlFlowTok{if}\NormalTok{(year }\SpecialCharTok{\%in\%}\NormalTok{ year\_A)\{}
\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{      dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
                    \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52A"}\NormalTok{) }\SpecialCharTok{\&}
                      \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\}}
  \ControlFlowTok{else}\NormalTok{\{}
\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{      dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
                    \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52D"}\NormalTok{) }\SpecialCharTok{\&}
                      \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\}}
\NormalTok{  temp }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979,}
                 \AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{,}
                 \AttributeTok{values\_to =} \StringTok{"hours\_work"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
\NormalTok{\}}
\CommentTok{\# list to save the iteration result}
\NormalTok{hours }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{()}
\CommentTok{\# getting the hours of work of all observations}
\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
\NormalTok{  hours[[ayear]] }\OtherTok{\textless{}{-}} \FunctionTok{get\_hour}\NormalTok{(ayear)}
\NormalTok{\}}
\CommentTok{\# unlist the hours of work}
\NormalTok{hours\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\SpecialCharTok{!!!}\NormalTok{hours)}
\end{Highlighting}
\end{Shaded}

The same way was also deployed to tidy the rate of wage by year and by ID. The difference is that the hourly rate has only one version of each year. The hours of work and the hourly rate are then joined to calculate the number of jobs that a respondent has and their mean hourly wage. Some observations have 0 in their hourly rate and work hours, which is considered an invalid value. Thus, these observations are set to be NA. Besides, there are some observations with unusual hours of work. Here, we also censored the observations with \textgreater{} 84 hours of work a week to be NA.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{get\_rate }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year) \{}
\NormalTok{  new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{    dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
                  \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\SpecialCharTok{\&}
                    \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year))) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979, }\AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"rate\_per\_hour"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_0"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
\NormalTok{\}}
\NormalTok{rates }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{()}
\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
\NormalTok{  rates[[ayear]] }\OtherTok{\textless{}{-}} \FunctionTok{get\_rate}\NormalTok{(ayear)}
\NormalTok{\}}
\NormalTok{rates\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\SpecialCharTok{!!!}\NormalTok{rates)}
\CommentTok{\# join hours and rates variable}
\NormalTok{hours\_wages }\OtherTok{\textless{}{-}} \FunctionTok{left\_join}\NormalTok{(rates\_all,}
\NormalTok{                         hours\_all,}
                         \AttributeTok{by =} \FunctionTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# set the 0 value in rate\_per\_hour as NA}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rate\_per\_hour =} \FunctionTok{ifelse}\NormalTok{(rate\_per\_hour }\SpecialCharTok{==} \DecValTok{0}\NormalTok{, }\ConstantTok{NA}\NormalTok{,}
\NormalTok{                                rate\_per\_hour),}
         \AttributeTok{hours\_work =} \FunctionTok{ifelse}\NormalTok{(hours\_work }\SpecialCharTok{==} \DecValTok{0} \SpecialCharTok{|}\NormalTok{ hours\_work }\SpecialCharTok{\textgreater{}} \DecValTok{84}\NormalTok{, }\ConstantTok{NA}\NormalTok{, hours\_work))}
\end{Highlighting}
\end{Shaded}

Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculated based on the availability of the \texttt{rate\_per\_hour} information. For example, the number of jobs of ID 1, based on \texttt{hours\_work}, is 2. However, since the information of hourly rate of \texttt{job\_02} is not available, the number of jobs is considered as 1.

Further, we calculated the mean hourly wage for each ID in each year using a weighted mean with the hours of work as the weight. However, there is a lot of missing value in \texttt{hours\_work} variable. In that case, we only calculated the mean hourly wage based on the arithmetic/regular mean method. Hence, we created a new variable to flag whether the mean hourly wage is weighted or regular. Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as arithmetic mean.

>>>>

@@

\if0\blind
{
  \title{\bf The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth}
  \author{
            Dewi Amaliah\thanks{Test}\\
      Department of Econometrics and Business Statistics, Monash University, \\
            Dianne Cook\\
      Department of Econometrics and Business Statistics, Monash University, \\
            Emi Tanaka\\
      Department of Econometrics and Business Statistics, Monash University, \\
            Kate Hyde\\
      Department of Econometrics and Business Statistics, Monash University, \\
            Nicholas Tierney\\
      , , }
        \maketitle
} \fi

@@
